# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
# Make a histogram of the total number of steps taken each day
steps_per_day <- aggregate(steps ~ date, data = activity, FUN = sum)
with(steps_per_day, hist(steps, breaks = nrow(steps_per_day), main = "Total Number of Steps Per Day", xlab = "Steps Per Day"))

# Calculate and report the mean and median total number of steps taken per day
mean_steps <- mean(steps_per_day$steps)
median_steps <- median(steps_per_day$steps)
```
The mean total number of steps taken per day is `r format(mean_steps)`

The median total number of steps taken per day is `r format(median_steps)`

## What is the average daily activity pattern?
```{r}
# Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
average_per_interval <- aggregate(steps ~ interval, data = activity, FUN = mean)
plot(average_per_interval, type = "l", xlab = "5-minute Interval", ylab = "Average # of Steps Taken", main = "Average Daily Activity Pattern")

# Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
max_steps_interval <- average_per_interval[which.max(average_per_interval$steps), "interval"]
```
The 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps is `r max_steps_interval`

## Imputing missing values
```{r}
# Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
total_missing <- sum(is.na(activity))
```
The total number of missing values in the dataset is `r total_missing`

```{r}
# Filling in all of the missing values in the dataset with the mean of the 5-minute interval averaged across all days
imputed <- numeric()
for(i in 1:nrow(activity)) {
        obs <- activity[i, ]
        if (is.na(obs$steps)) {
                steps <- subset(average_per_interval, interval == obs$interval)$steps
        } else {
                steps <- obs$steps
        }
        imputed <- c(imputed, steps)
}

# Create a new dataset that is equal to the original dataset but with the missing data filled in
imputed_activity <- activity
imputed_activity$steps <- imputed

# Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
imputed_steps_per_day <- aggregate(steps ~ date, data = imputed_activity, FUN = sum)

hist(imputed_steps_per_day$steps, breaks = nrow(imputed_steps_per_day), xlab = "Steps Per Day", main = "Total Number of Steps Per Day")

imputed_mean_steps <- mean(imputed_steps_per_day$steps)
imputed_median_steps <- median(imputed_steps_per_day$steps)
```
The mean total number of steps taken per day is `r format(imputed_mean_steps)`

The median total number of steps taken per day is `r format(imputed_median_steps)`

After imputing missing data, mean is unchanged, while median is a little bit increase.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
# Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
wd <- strptime(imputed_activity$date, "%Y-%m-%d")
wd <- weekdays(wd)
l <- wd == "Sunday" | wd == "Saturday"
wd[l] <- "weekend"
wd[!l] <- "weekday"
wd <- as.factor(wd)
imputed_activity$daytype <- wd

# Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
require(plyr)
average_steps <- ddply(imputed_activity, .(interval, daytype), summarize, steps = mean(steps))

require(lattice)
xyplot(steps ~ interval | daytype, data = average_steps, layout = c(1, 2), type = "l", xlab = "Interval", ylab = "Number of Steps")
```
